---
title: "Vignette for ridgereg"
author: "Yumeng Li, Mattias Karlsson, Ashraf Sarhan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Vignette Goal
This vignette guides you through how to predict observations using the ridgereg() function from package Linreg. In this case we will predict flight delays within JFK airport.

## Function Discription
Ridgereg(formula, data, lambda) 

Ridge regression can be a good alternative when we have a lot of covariates (when p > n) or in the situation of multicollinearity.

## Data
Read in the weather dataset and the flights dataset from the nycflights13 package and remove eventual variables you do not believe to have a predictive value.
```
library(nycflights13)
library(dplyr)
data(weather)
data(flights)

# Pick needed variables
delay_d <- select(flights,time_hour,origin,dep_delay)
delay_d <- filter(delay_d,dep_delay>0)

# Subset JFK airport
delay_d_jfk <- filter(delay_d,origin == "JFK")
```
Variable dep_delay say how many minutes the plane was delayed. date_hour variable is the variable that is used to match the weather data. Splits the data so dep_delays contains only positive values, because we are only interested in flight that has been delayed.
```{r, echo = FALSE}
suppressMessages(library(nycflights13))
suppressMessages(library(dplyr))
suppressMessages(library(linreg))
#test
data(weather)
data(flights)

# Pick needed variables
delay_d <- select(flights,time_hour,origin,dep_delay)
delay_d <- filter(delay_d,dep_delay>0)
# Subset JFK airport
delay_d_jfk <- filter(delay_d,origin == "JFK")
head(delay_d_jfk)
```

```
# Weather data
weather_d <- select(weather,time_hour,origin, wind_speed,precip,visib,month)

# Subset JFK airport
weather_d_jfk <- filter(weather_d, origin == "JFK")

# Match datasets by time_hour variable

delay_d_complete <- inner_join(delay_d_jfk, weather_d_jfk, by = "time_hour")
delay_d_complete <- select(delay_d_complete,-origin.y,-origin.x)

# Keep row that have a positive value on dep_delay
delay_d_complete <- filter(delay_d_complete,dep_delay>0)

# Convert to data.frame lets us use createDataPartition function
delay_d_complete <- as.data.frame(delay_d_complete)
delay_d_complete$month <- as.factor(delay_d_complete$month)
```
Weather data set contains the explanatory variables to be used in the model.

precip = Preciptation, in inches
wind_speed = Wind speed (in mph)
visib = Visibility in miles
month = Time of recording

```{r,echo = FALSE}
# Weather data
weather_d <- select(weather,time_hour,origin, wind_speed,precip,visib,month)

# Subset JFK airport
weather_d_jfk <- filter(weather_d, origin == "JFK")

# Match datasets by time_hour variable

delay_d_complete <- inner_join(delay_d_jfk, weather_d_jfk, by = "time_hour")
delay_d_complete <- select(delay_d_complete,-origin.y,-origin.x)

# Keep row that have a positive value on dep_delay
#delay_d_complete <- filter(delay_d_complete,dep_delay>0)

delay_d_complete <- as.data.frame(delay_d_complete)
delay_d_complete$month <- as.factor(delay_d_complete$month)

head(delay_d_complete)
```

## Test/Train/Validation dataset
We will now divide the data set into three parts, testing, training and validation sets. The division of the observations to each set is, 5%, 80% and 15%
Each set is
```
library(lattice)
library(ggplot2)
library(caret)

set.seed(123456789)
test_rows <- createDataPartition(y = delay_d_complete$month,
                                 p = 0.05,
                                 list = FALSE)
test_sample <- delay_d_complete[test_rows,]

# Remove observation that got into the test sample
delay_d_complete <- delay_d_complete[-test_rows,]

train_rows <- createDataPartition(y = delay_d_complete$month,
                                  p = 0.80,
                                  list = FALSE)
train_sample <- delay_d_complete[train_rows,]

# Remove observation from that got into train sample

delay_d_complete <- delay_d_complete[-train_rows,]

validation_rows <- createDataPartition(y = delay_d_complete$month,
                                          p = 0.15,
                                          list = FALSE)
validation_sample <- delay_d_complete[validation_rows,]

```
```{r, echo = FALSE}
suppressMessages(library(lattice))
suppressMessages(library(ggplot2))
suppressMessages(library(caret))

set.seed(123456789)
test_rows <- createDataPartition(y = delay_d_complete$month,
                                 p = 0.05,
                                 list = FALSE)
test_sample <- delay_d_complete[test_rows,]

```
Test dataset
```{r,echo=FALSE}
head(test_sample)
```

Train dataset
```{r, echo=FALSE}
# Remove observation that got into the test sample
delay_d_complete <- delay_d_complete[-test_rows,]

train_rows <- createDataPartition(y = delay_d_complete$month,
                                  p = 0.80,
                                  list = FALSE)
train_sample <- delay_d_complete[train_rows,]

head(train_sample)

```
Validation dataset
```{r, echo = FALSE}
# Remove observation from that got into train sample

delay_d_complete <- delay_d_complete[-train_rows,]

validation_rows <- createDataPartition(y = delay_d_complete$month,
                                          p = 0.15,
                                          list = FALSE)
validation_sample <- delay_d_complete[validation_rows,]

head(validation_sample)
```

# Ridge Regression - Validation dataset
Train ridge regressions models for di↵erent values of lambda and evaluate the root mean squared error (RMSE)on the validation set.
In this case, we check the lambda values between 0-10.
```{r, echo = TRUE}
train_f <- function(data,lambda,formula){
  
  count <- seq(from = 0,to = lambda, by = 0.05)
  save <- data.frame("lambda" = count)
  y <- as.character(formula)[2]
  
  for(i in 1:length(count)){
    call <- ridgereg(formula,data,lambda = count[i])
    save$RMSE[i] <- sqrt(mean((data[-1,y] - call$fitted_values)^2))
  }
  return(save[which.min(save$RMSE),])
}

call_f <- formula(dep_delay ~ wind_speed+precip+visib)
train_f(data = validation_sample,lambda = 10,formula = call_f)

```

# Ridge Regression - Test dataset
We now predict the observations for the test set by using the evaluated lambda value above.
```{r, echo = TRUE}
prediction <- ridgereg(data = test_sample,lambda = 0,formula = call_f)
head(prediction$fitted_values)
```
The RMSE value for the model is presented below.
```
#### Include ridgereb in caret package
svmGrid <- function(x, y, len = NULL, search = "grid") {
  #library(kernlab)
  ## This produces low, middle and high values for sigma 
  ## (i.e. a vector with 3 elements). 
  #sigmas <- sigest(as.matrix(x), na.action = na.omit, scaled = TRUE)  
  ## To use grid search:
  if(search == "grid") {
    out <- expand.grid(lambda = c(0, 10^seq(-1, -4, length = len - 1)))
  } else {
    ## For random search, define ranges for the parameters then
    ## generate random values for them
    out <- data.frame(lambda = 10^runif(len, min = -5, 1))
  }
  out
}

train_x <- validation_sample
train_y <- validation_sample$dep_delay
train_p <- "lambda"
train_y <- 5

svmFit <- function(x,y,param){
  ksvm(as.matrix(x),
       y = y,
       kernel = rbfdot,
       kpar = list(lambda = param$lambda))
}

ridge_list <- list(library = "linreg",
                   type = "regression",
                   parameters = data.frame(parameter = "lambda",
                                           class = "numeric",
                                           label = "Lambda"),
                   grid = svmGrid,
                   #(x = train_x,y = train_y, param = train_p)
                   fit = svmFit)

Lambda_optimize <- train(data = validation_sample,
      x = validation_sample[1:100,3:5],
      y = c(1:100),
      form = call_f,
      type = ridge_list,
      maxit = 100)
```


